Type-Safe Modular Parsing

### 2017-01-07 paper/sections/CaseStudy.tex
Some changes

To discover the reasons of slower execution time, we made some more experiments
on two possible factors which could affect the performance.
They are object algebra and the longest matching alternative combinator.
We use object algebra for ASTs and the longest matching alternative combinator \inlinecode{|||} for parsing,
while Ilya's implementation uses case class and the ordinary alternative combinator.

We implemented two more versions. One is a modified version of our implementation,
with object algebra replaced by case class for the ASTs.
The other is a modified version of Ilya's implementation, using the longest matching
alternative combinator instead of the ordinary one.

@@

Table \ref{tab:ext-comparison} shows part of the comparison results of these four versions,
full table is available online []. It suggests that the difference of running time between
using object algebra and class is little, roughly 1\%.
The usage of longest matching combinator slows the performance by 7\%. The main reason of slower
execution time may be the overall structure of the modular parsing approach, because we indeed have
more intermediate function calls and method overriding. However, it is worth metioning that
because of the memoization technique of Packrat parsers, we are only constant times
slower, the algorithmic complexity is still the same.

>>>>

To discover the reasons of slower execution time, we made some more experiments
on two possible factors which could affect the performance.
They are object algebra and the longest match alternative combinator.
We use object algebra for ASTs and the longest match alternative combinator \inlinecode{|||} for parsing,
while Ilya's implementation uses case class and the ordinary alternative combinator.

We implemented two more versions. One is a modified version of our implementation,
with object algebra replaced by case class for the ASTs.
The other is a modified version of Ilya's implementation, using the longest match
alternative combinator instead of the ordinary one.

@@

Table \ref{tab:ext-comparison} shows part of the comparison results of these four versions,
full table is available online []. It suggests that the difference of running time between
using object algebra and class is little, roughly 1\%.
The usage of longest match combinator slows the performance by 7\%. The main reason of slower
execution time may be the overall structure of the modular parsing approach, because we indeed have
more intermediate function calls and method overriding. However, it is worth metioning that
because of the memoization technique of Packrat parsers, we are only constant times
slower, the algorithmic complexity is still the same.

### 2017-01-07 paper/sections/Discussion.tex
Some changes

\lstinputlisting[linerange=139-144]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
The above parser shows part of the code, for the space reason we omit details. With OO inheritance, we can actually override \lstinline{pLam}
in a subtype of \lstinline{BaseParser}, and give a different implementation. It will make a difference during runtime due to dynamic
dispatch.

\lstinputlisting[linerange=148-154]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
On the other hand, the lexer can also be updated, since keywords and delimiters are represented by sets of strings.
One can even ``eliminate'' a production rule in the extension, by actually overriding it with a failure parser.

@@

\lstinputlisting[linerange=158-164]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=NEWPARSER_UNTYPEDLAM

>>>>

\lstinputlisting[linerange=106-111]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
The above parser shows part of the code, for the space reason we omit details. With OO inheritance, we can actually override \lstinline{pLam}
in a subtype of \lstinline{BaseParser}, and give a different implementation. It will make a difference during runtime due to dynamic
dispatch.

\lstinputlisting[linerange=115-121]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
On the other hand, the lexer can also be updated, since keywords and delimiters are represented by sets of strings.
One can even ``eliminate'' a production rule in the extension, by actually overriding it with a failure parser.

@@

\lstinputlisting[linerange=125-131]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=NEWPARSER_UNTYPEDLAM

### 2017-01-07 paper/sections/Inheritance.tex
Some changes

\begin{itemize}
\item \textbf{Extension 1}: adding a new case (or rather, a new construct) to \lstinline{Expr};
\item \textbf{Extension 2}: adding a new operation to \lstinline{Expr}.
\end{itemize}

@@

\lstinputlisting[linerange=98-100]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_LAM

And since we already have the parser for literals and additions, we would
like to build the new parser by reusing the old one. Here one may quickly come up
with the following attempt, where a new parser is defined for \lstinline{Var}, then composed
with \lstinline{pExpr}:

\lstinputlisting[linerange=62-63]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=BAD_ATTEMPT

Unfortunately, it fails to parse some expressions like \lstinline{"1 + x"}, which is obviously valid in the new grammar.
The reason is that \lstinline{pAdd} makes two recursive calls to parse sub-expressions, by using \lstinline{pExpr}, which
covers both cases in the old grammar. But the newly added case \lstinline{pVar} is not observed by the recursive \lstinline{pExpr},
hence the parser does not work as expected. Still, one may continue to change the recursive call in \lstinline{pAdd} with \lstinline{pVarExpr},
but modification on existing code sacrifices separate compilation, as mentioned by the Expression Problem.

@@

@@

\lstinputlisting[linerange=68-73]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=INHERITANCE_APPROACH
Now \lstinline{VarExprParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
method overriding in inheritance. When the input \lstinline{"1 + x"} is fed to the parser \lstinline{this.pExpr}, it firstly delegates
the work to \lstinline{super.pExpr}, which parses literals and additions. However, the recursive call \lstinline{pExpr} in \lstinline{pAdd}
actually refers to \lstinline{this.pExpr} again, due to dynamic dispatch. It implies that the recursive call has included the extension. This
elegant inheritance-based approach opens the recursion of parsers in a modular way.

Notice that Scala has a linearized-style multiple inheritance for traits. It can be very helpful when composing several languages. Suppose now
we want to compose the parsers for expressions from pre-defined languages \lstinline{LanguageA} and \lstinline{LanguageB} using alternative.
The use of keyword \lstinline{super} is able to
specify the implementation from inheritance, like:

\lstinputlisting[linerange=78-84]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=MULTIPLE_INHERITANCE

>>>>

\begin{itemize}
\item \textbf{Extension 1}: adding a new case (or rather, a new construct) to \inlinecode{Expr};
\item \textbf{Extension 2}: adding a new operation to \inlinecode{Expr}.
\end{itemize}

@@

\lstinputlisting[linerange=30-32]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_SIMPLE_LAM

And since we already have the parser for literals and additions, we would
like to build the new parser by reusing the old one. Here one may quickly come up
with the following attempt, where a new parser is defined for \inlinecode{Var}, then composed
with \inlinecode{pExpr}:

\huang{Extending from old parser is necessory to get the `pExpr', added a surrounding trait.}
\lstinputlisting[linerange=43-46]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_BAD_ATTEMPT

Unfortunately, it fails to parse some expressions like \inlinecode{"1 + x"}, which is obviously valid in the new grammar.
The reason is that \inlinecode{pAdd} makes two recursive calls to parse sub-expressions, by using \inlinecode{pExpr}, which
covers both cases in the old grammar. But the newly added case \inlinecode{pVar} is not observed by the recursive \inlinecode{pExpr},
hence the parser does not work as expected. Still, one may continue to change the recursive call in \inlinecode{pAdd} with \inlinecode{pVarExpr},
but modification on existing code sacrifices separate compilation, as mentioned by the Expression Problem.

@@

\huang{It seems a little bit weird since we were already using inheritance in the `bad attempt'. I think we could focus on overriding here}

@@

\lstinputlisting[linerange=50-55]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH
Now \inlinecode{VarExprParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
method overriding in inheritance. When the input \inlinecode{"1 + x"} is fed to the parser \inlinecode{this.pExpr}, it firstly delegates
the work to \inlinecode{super.pExpr}, which parses literals and additions. However, the recursive call \inlinecode{pExpr} in \inlinecode{pAdd}
actually refers to \inlinecode{this.pExpr} again, due to dynamic dispatch. It implies that the recursive call has included the extension. This
elegant inheritance-based approach opens the recursion of parsers in a modular way.

Notice that Scala has a linearized-style multiple inheritance for traits. It can be very helpful when composing several languages. Suppose now
we want to compose the parsers for expressions from pre-defined languages \inlinecode{LanguageA} and \inlinecode{LanguageB} using alternative.
The use of keyword \inlinecode{super} is able to
specify the implementation from inheritance, like:

\lstinputlisting[linerange=60-66]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE

### 2017-01-07 paper/sections/ObjectAlgebras.tex
Some changes

The inheritance-based approach benefits us from building extensible parsers, based on an OO class hierarchy.
Nevertheless, the addition of new operations over the AST becomes problematic. Often it requires modification on
the existing code, which breaks modularity. In this section, we will have a closer look at how it motivates us to
use Object Algebras for full extensibility.

@@

Now it comes to extension 2 in Section~\ref{subsec:dimensions}. In the previous code, we have presented pretty-printing as
an example of operations on the AST. Suppose now we also want to collect free variables on the expression language, which models
literals, additions and variables. Based on the traditional OO class hierarchy, one attempt would be extending \lstinline{Expr} with the new
operation directly:

\lstinputlisting[linerange=89-89]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=ATTEMPT_EXPRWITHFREEVARS
It refers to one Scala solution to the Expression Problem []. But then programmers have to define new classes for literals, additions and variables
by extending the old ones together with \lstinline{FreeVarsExpr}. The critical issue is that, the existing parsing code only generates objects of the old
types, again it is a modularity issue, since it is a bad idea to modify existing code again and again with extensions. Other solutions including modelling constructors as functions introduce unnecessary complexity.

@@

\lstinputlisting[linerange=8-11]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_ALG
The language has two constructors: literals and additions. The trait \lstinline{ExprAlg} is called an \textit{Object Algebra interface},
whereas as a factory, it cannot produces objects for the expressions directly like traditional approaches, but abstract the results in its
type parameter \lstinline{E} instead. To realize an operation on expressions, we simply instantiate the type parameter by a concrete type and
provides implementations for all cases. Below is an example of pretty-printing an expression:

\lstinputlisting[linerange=15-18]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
Here \lstinline{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
Furthermore, one can define a new trait to implement evaluation on arithmetic expressions:

\lstinputlisting[linerange=22-25]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EVAL

@@

\lstinputlisting[linerange=29-31]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
Now pretty-printing on the new language can be realized without modifying existing code:

\lstinputlisting[linerange=35-37]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
We observe that only the new case is implemented for pretty-printing, and the others have been inherited.
It offers programmers with great code reuse, in the sense that existing cases can also be
overridden in \lstinline{VarExprPrint} using \lstinline{"override def"}.

@@

\huang{todo: better name for build}\haoyuan{makeExp}
\lstinputlisting[linerange=41-41]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
Such a method implicitly represents the expression \lstinline{"1 + x"}. The code

@@

\lstinputlisting[linerange=45-56]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
Here the \lstinline{VarExprAlg} expression \lstinline{"1 + x"} is changed to \lstinline{ExprAlg} expression \lstinline{"1 + 2"}, before the pretty-printing algebra is applied. The abstract method
\lstinline{alg} in \lstinline{Refactor} can actually be any algebra simply for delegation. By passing a concrete implementation
to \lstinline{alg}, \lstinline{Refactor} will be able to transform a \lstinline{VarExprAlg} expression before we obtain the result.

@@

\lstinputlisting[linerange=60-68]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
Such a parser works fine, yet it does not appear to be extensible. For example, we have demonstrated in Section~\ref{subsec:inheritance-approach} that method overriding can update \lstinline{pExpr} for an extended syntax, but now \lstinline{pExpr} will have a different type, like \lstinline{VarExprAlg[E] =>} \lstinline{Parser[E]} where the difference appears in the argument. The difference of types prohibits method overriding, however, this problem refers to \textit{binary methods}. \haoyuan{Shall we mention this?}

One solution to this problem is taking out the argument as a single field, hence it supports overriding as well. Therefore we refactor the parsing code as follows:
\lstinputlisting[linerange=72-81]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASE_OA_PARSER
This is precisely the pattern that we advocate for modular parsing. Note that \lstinline{pE} is the one that is used recursively, exported by the trait and considered to be overridden from inheritance. The reason why we use an extra and seemingly redundant \lstinline{pExprE}, is that \lstinline{pE} can no longer be referred to by the \lstinline{super} keyword, as a limitation of Scala for generics, hence we can directly invoke \lstinline{pExprE} as the specific \lstinline{pE} in \lstinline{ExprOAParser}. And we still want code reuse for the composition of \lstinline{pLit} and \lstinline{pAdd}. Now it is worth trying on this pattern with an extension, like \lstinline{VarExprAlg} which we have been using:
\lstinputlisting[linerange=85-96]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=EXT_OA_PARSER
In the client code shown above, an instance of algebra is fed to \lstinline{VarExprOAParser} for pretty-printing expressions. Then the trait exports \lstinline{pE}, parses the input and applies the algebra. The result shows that such a pattern is a nice solution to modular parsing. At this point, one may argue that with a different operation like evaluation, the redundant code is required to obtain the parser, and it results in parsing the same input twice. This is generally solved in the original Object Algebras paper [] by composing two algebras into one in advance.

@@

\lstinputlisting[linerange=100-129]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
Besides lexing, the trait \lstinline{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. For the same reason,
we use \lstinline{pTypedLamT} and \lstinline{pTypedLamE} as copies of current \lstinline{pT} and \lstinline{pE}, recursively. \lstinline{pT} and \lstinline{pE} are actually the parsers for recursion.

@@

\lstinputlisting[linerange=133-135]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
Pretty-printing the parsed results just works as expected.

>>>>

The inheritance-based approach benefits us from building extensible parsers, based on an OO class hierarchy.
Nevertheless, the addition of new operations over the ASTs becomes problematic. Often it requires modification on
the existing code, which breaks modularity. In this section, we will have a closer look at how it motivates us to
use Object Algebras for full extensibility.

@@

Now it comes to extension 2 in Section~\ref{subsec:dimensions}. In the previous code, we have presented pretty-printing as
an example of operations on the ASTs. Suppose now we also want to collect free variables on the expression language, which models
literals, additions and variables. Based on the traditional OO class hierarchy, one attempt would be extending \lstinline{Expr} with the new
operation directly:

\lstinputlisting[linerange=19-19]{../Scala/Parser/src/PaperCode/Sec4OA/Code1.scala}% APPLY:linerange=ATTEMPT_EXPRWITHFREEVARS
It refers to one Scala solution to the Expression Problem []. But then programmers have to define new classes for literals, additions and variables
by extending the old ones together with \lstinline{FreeVarsExpr}. The critical issue is that, the existing parsing code only generates objects of the old
types, again it is a modularity issue, since it is a bad idea to modify existing code again and again with extensions. Other solutions including modelling constructors as functions introduce unnecessary complexity.

@@

\lstinputlisting[linerange=7-10]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
The language has two constructors: literals and additions. The trait \lstinline{ExprAlg} is called an \textit{Object Algebra interface},
whereas as a factory, it cannot produces objects for the expressions directly like traditional approaches, but abstract the results in its
type parameter \lstinline{E} instead. To realize an operation on expressions, we simply instantiate the type parameter by a concrete type and
provides implementations for all cases. Below is an example of pretty-printing an expression:

\lstinputlisting[linerange=14-17]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
Here \lstinline{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
Furthermore, one can define a new trait to implement evaluation on arithmetic expressions:

\lstinputlisting[linerange=21-24]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL

@@

\lstinputlisting[linerange=28-30]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
Now pretty-printing on the new language can be realized without modifying existing code:

\lstinputlisting[linerange=34-36]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
We observe that only the new case is implemented for pretty-printing, and the others have been inherited.
It offers programmers with great code reuse, in the sense that existing cases can also be
overridden in \lstinline{VarExprPrint} using \lstinline{"override def"}.

@@

\huang{todo: better name for build}\haoyuan{makeExp}
\lstinputlisting[linerange=8-8]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
Such a method implicitly represents the expression \lstinline{"1 + x"}. The code

@@

\lstinputlisting[linerange=12-23]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
Here the \lstinline{VarExprAlg} expression \lstinline{"1 + x"} is changed to \lstinline{ExprAlg} expression \lstinline{"1 + 2"}, before the pretty-printing algebra is applied. The abstract method
\lstinline{alg} in \lstinline{Refactor} can actually be any algebra simply for delegation. By passing a concrete implementation
to \lstinline{alg}, \lstinline{Refactor} will be able to transform a \lstinline{VarExprAlg} expression before we obtain the result.

@@

\lstinputlisting[linerange=27-35]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
Such a parser works fine, yet it does not appear to be extensible. For example, we have demonstrated in Section~\ref{subsec:inheritance-approach} that method overriding can update \lstinline{pExpr} for an extended syntax, but now \lstinline{pExpr} will have a different type, like \lstinline{VarExprAlg[E] =>} \lstinline{Parser[E]} where the difference appears in the argument. The difference of types prohibits method overriding, however, this problem refers to \textit{binary methods}. \haoyuan{Shall we mention this?}

One solution to this problem is taking out the argument as a single field, hence it supports overriding as well. Therefore we refactor the parsing code as follows:
\lstinputlisting[linerange=39-48]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASE_OA_PARSER
This is precisely the pattern that we advocate for modular parsing. Note that \lstinline{pE} is the one that is used recursively, exported by the trait and considered to be overridden from inheritance. The reason why we use an extra and seemingly redundant \lstinline{pExprE}, is that \lstinline{pE} can no longer be referred to by the \lstinline{super} keyword, as a limitation of Scala for generics, hence we can directly invoke \lstinline{pExprE} as the specific \lstinline{pE} in \lstinline{ExprOAParser}. And we still want code reuse for the composition of \lstinline{pLit} and \lstinline{pAdd}. Now it is worth trying on this pattern with an extension, like \lstinline{VarExprAlg} which we have been using:
\lstinputlisting[linerange=52-63]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=EXT_OA_PARSER
In the client code shown above, an instance of algebra is fed to \lstinline{VarExprOAParser} for pretty-printing expressions. Then the trait exports \lstinline{pE}, parses the input and applies the algebra. The result shows that such a pattern is a nice solution to modular parsing. At this point, one may argue that with a different operation like evaluation, the redundant code is required to obtain the parser, and it results in parsing the same input twice. This is generally solved in the original Object Algebras paper [] by composing two algebras into one in advance.

@@

\lstinputlisting[linerange=67-96]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
Besides lexing, the trait \lstinline{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. For the same reason,
we use \lstinline{pTypedLamT} and \lstinline{pTypedLamE} as copies of current \lstinline{pT} and \lstinline{pE}, recursively. \lstinline{pT} and \lstinline{pE} are actually the parsers for recursion.

@@

\lstinputlisting[linerange=100-102]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
Pretty-printing the parsed results just works as expected.

### 2017-01-07 paper/sections/PackratParsing.tex
Some changes

This section discusses the algorithmic challenges introduced by modular parsing and argues that parser combinators from Packrat Parsing
are suitable to address those challenges.

@@

\subsection{Packrat Parsing}\label{subsec:packratparsing}
Fortunately some more advanced parsing techniques such as Packrat parsing~\cite{Ford2002}, address the limitations of simple parser combinators such as Parsec. Packrat parsers use memoization to record the result of applying each parser at each position of the input, so that repeated computation is eliminated.  Moreover, theoretically the algorithm behind Packrat parsers supports both direct and indirect left-recursion~\cite{warth2008}. All of these properties are very suitable for modularity, thus we decided to use Packrat parsers as the underlying parsing technique in our modular parsing library.
\begin{comment}
It is worth mentioning that the choice of parser combinators will not
affect the other parts of our library. One can choose other parser
combinators like Parsec, in cases that the performance and supporting
of left-recursion are not major concerns. A different library can even build a new
\name with fancy features or higher efficiency.
\end{comment}
Scala has a standard parser combinator library\footnote{https://github.com/scala/scala-parser-combinators}
which implements Packrat parsers.
The library provides a number of parser combinators, together with some basic parsers.
Below we present an example to illustrate Scala parsers.
%
%\bruno{So, we make a big fuss identifying the problems with other
%parser combinators in the previous subsection and now we present
%an example that does not suffer from any of the previous problems!
%What makes sense here is an example that uses: a \emph{left-recursive grammar}; the \emph{longest match combinator}; and where we can argue that
%there's no need for the user to do manual backtracking. Just use
%the example with the left-recursive grammar for parsing integers and additions and argue that none of the limitations
%discussed previously applies!
%}

@@

\huang{doesn't compile}

\begin{figure}[t]
\centering
\lstinputlisting[linerange=7-27]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_EXPR
\caption{Packrat Parsing for an arithmetic language.}\label{fig:packrat-arith}
\end{figure}

In the trait \lstinline{ExprParser}, \lstinline{lexical} is used for lexing. \lstinline{pLit} can parse an integer for \lstinline{Lit},
whereas \lstinline{pAdd} handles addition and creates an object of \lstinline{Add}. It parses two sub-expressions by calling \lstinline{pExpr}
recursively. Finally \lstinline{pExpr} composes \lstinline{pLit} and \lstinline{pAdd} using the alternative combinator with longest match.
Table~\ref{tab:packrat} gives a closer look at part of the API that we have used in our work.

@@

Finally, a generic \inlinecode{parse} function is used for testing:

\lstinputlisting[linerange=52-57]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=PACKRAT_RUNPARSER

It is worth mentioning that such a grammar is left-recursive, yet well supported by Packrat parsers. It also employs automatic
longest match composition by using the library function \lstinline{|||}. Finally, Packrat parsers do not suffer from the backtracking
problem, as the technique uses memoization to record intermediate parsing results. We will use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} in the rest of paper.

>>>>

This section discusses the algorithmic challenges introduced by modular parsing and argues that parser combinators from Packrat parsing
are suitable to address those challenges.

@@

\subsection{Packrat Parsing}\label{subsec:packratparsing}
Fortunately some more advanced parsing techniques such as Packrat parsing~\cite{Ford2002}, address the limitations of simple parser combinators such as Parsec. Packrat parsers use memoization to record the result of applying each parser at each position of the input, so that repeated computation is eliminated.  Moreover, theoretically the algorithm behind Packrat parsers supports both direct and indirect left-recursion~\cite{warth2008}. All of these properties are very suitable for modularity, thus we decided to use Packrat parsers as the underlying parsing technique in our modular parsing library.
\begin{comment}
It is worth mentioning that the choice of parser combinators will not
affect the other parts of our library. One can choose other parser
combinators like Parsec, in cases that the performance and supporting
of left-recursion are not major concerns. A different library can even build a new
\name with fancy features or higher efficiency.
\end{comment}
Scala has a standard parser combinator library\footnote{https://github.com/scala/scala-parser-combinators}
\cite{moors2008parser} which implements Packrat parsers.
The library provides a number of parser combinators, together with some basic parsers.
Below we present an example to illustrate Scala parsers.
%
%\bruno{So, we make a big fuss identifying the problems with other
%parser combinators in the previous subsection and now we present
%an example that does not suffer from any of the previous problems!
%What makes sense here is an example that uses: a \emph{left-recursive grammar}; the \emph{longest match combinator}; and where we can argue that
%there's no need for the user to do manual backtracking. Just use
%the example with the left-recursive grammar for parsing integers and additions and argue that none of the limitations
%discussed previously applies!
%}

@@

\begin{figure}[t]
\centering
\lstinputlisting[linerange=6-29]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
\caption{Packrat Parsing for an arithmetic language.}\label{fig:packrat-arith}
\end{figure}

In the trait \lstinline{ExprParser}, \lstinline{lexical} is used for lexing. \lstinline{pLit} can parse an integer for \lstinline{Lit},
whereas \lstinline{pAdd} handles addition and creates an object of \lstinline{Add}. It parses two sub-expressions by calling \lstinline{pExpr}
recursively. Finally \lstinline{pExpr} composes \lstinline{pLit} and \lstinline{pAdd} using the longest match alternative combinator \inlinecode{|||}.
Table~\ref{tab:packrat} gives a closer look at part of the API that we have used in our work.

@@

It is worth mentioning that such a grammar is left-recursive, yet well supported by Packrat parsers. It also employs automatic
longest match composition by using the combinator \lstinline{|||} in the library. Furthermore, Packrat parsers do not suffer from the backtracking
problem, as the technique uses memoization to record intermediate parsing results.

To shorten our code, we will use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} and a generic \inlinecode{parse} function for testing in the rest of paper. The last line demonstrates how we parse a valid expression \inlinecode{1 + 2} using our parser.

\lstinputlisting[linerange=36-43]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_RUNPARSER
